# EDA on Body Performance

## \<\>. Loading Library in the code

```{r}
library(ggplot2)
library(gridExtra)
library(corrplot)
library(tidyr)
library(e1071)
```

## \<\>. Loading the Data

```{r}
setwd(getwd())
main_data<-read.csv("bodyPerformance.csv")
View(main_data)
```

## \<\>. How many Male and Female are in the data?

```{r}
################################################################################
# Male and female %
################################################################################
gender_count <- table(main_data$gender)
data <- data.frame(
  Category = c("Female", "Male"),
  Value = c(gender_count)
)
print(gender_count)
data$Percentage <- (data$Value / sum(data$Value)) * 100
p <- ggplot(data, aes(x = "", y = Value, fill = Category)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "Pie Chart of Percentage of male and female") +
  geom_text(aes(label = paste(round(Percentage,2), "%")),
            position = position_stack(vjust = 0.5)) +
  theme_void()
print(p)
```

-Male : 63.22% (8467 Males)

-Female: 36.78% (4926 Females)

## \<\>. How many number of Male and female are in each Class?

```{r}
################################################################################
# no of male and female in each class//bar
################################################################################

a<-table(main_data$gender,main_data$class)
data <- data.frame(
  Category = c("A", "B", "C", "D"),
  Female = c(a[1],a[3],a[5],a[7]),
  Male = c(a[2],a[4],a[6],a[8])
)
data_long <- gather(data, key = "Group", value = "Value", -Category)
p <- ggplot(data_long, aes(x = Category, y = Value, fill = Group)) +
  geom_bar(stat = "identity") +
  labs(title = "Bar Plot Between No. of people and Class", y = "No. of People",x="Class") +
  geom_text(aes(label = Value), vjust = -0.5, position = position_dodge(width = 0.9), size = 3) +
  scale_fill_manual(values = c("Female" = "salmon", "Male" = "skyblue"))
print(p)
```

-A : Most number of females and Least number of Males

-C : Most number of Males and Least number of Females

-All class have more numbers of male than female

## \<\>. Distribution of Age in the data

```{r}
################################################################################
# frequency of age//histogram
################################################################################
data <- data.frame(Values = main_data$age)
age_skew<-e1071::skewness(main_data$age)
cat("Skewness of Age is ",age_skew)
ggplot(data, aes(x = Values)) +
  geom_histogram(binwidth = 8, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Age", x = "Age(Years)", y = "Frequency") + 
  theme_minimal() 

```

-Fewer people with higher age

-more people with lower age

-Age is positively Skewed so "Median" is lower than the "Mean"

## \<\>. How age is Distributed in the each Class?

```{r}
################################################################################
# box plot between age and class
################################################################################


data <- data.frame(
  Category = main_data$class,
  Value = main_data$age
)
p <- ggplot(data, aes(x = Category, y = Value, fill = Category)) +
  geom_boxplot(width = 0.2) +
  scale_fill_manual(values = c("orange", "green", "skyblue", "purple")) +
  coord_flip()+
  # theme_minimal() +
  labs(title = "Horizontal Box Plots Between Class and Age",x='Class',y='Age')
print(p)



################################################################################
# Bar plot between age and class
################################################################################

A_mean_age <- mean(main_data$age[main_data$class == 'A'])
B_mean_age <- mean(main_data$age[main_data$class == 'B'])
C_mean_age <- mean(main_data$age[main_data$class == 'C'])
D_mean_age <- mean(main_data$age[main_data$class == 'D'])

data4 <- data.frame(Category = c('A','B','C','D'), Value = c(A_mean_age, B_mean_age, C_mean_age,D_mean_age))

p4 <- ggplot(data4, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust =-0.1, size = 4) +
  labs(title = "", x = "Class", y = "Mean Age")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Age")
p4

Aq1 <- quantile(main_data$age[main_data$class == 'A'], 0.25)
Aq3 <- quantile(main_data$age[main_data$class == 'A'], 0.75)
cat("Q1 of A:", Aq1, "\n")
cat("Q3 of A:", Aq3, "\n")

Bq1 <- quantile(main_data$age[main_data$class == 'B'], 0.25)
Bq3 <- quantile(main_data$age[main_data$class == 'B'], 0.75)
cat("Q1 of B:", Bq1, "\n")
cat("Q3 of B:", Bq3, "\n")

Cq1 <- quantile(main_data$age[main_data$class == 'C'], 0.25)
Cq3 <- quantile(main_data$age[main_data$class == 'C'], 0.75)
cat("Q1 of C:", Cq1, "\n")
cat("Q3 of C:", Cq3, "\n")

Dq1 <- quantile(main_data$age[main_data$class == 'D'], 0.25)
Dq3 <- quantile(main_data$age[main_data$class == 'D'], 0.75)
cat("Q1 of D:", Dq1, "\n")
cat("Q3 of D:", Dq3, "\n")
```

every class contains people from every age group but

\- A has a group of people which belongs to 25 to 44 age group and

-D has a group of people which belongs to 25 to 50 age group

-A has lesser median age than other class so better performance

-D has higher median age than other class so better performance

change in median age is very low , so it is

-more or less each class have same distribution of age , so we try to find how body and physical attributes classify them in 4 class

## \<\>. Is there any category where Female performed better than male?

```{r}
################################################################################
# mean of PA and gender
################################################################################

M_mean_grip <- mean(main_data$gripForce[main_data$gender == 'M'])
F_mean_grip <- mean(main_data$gripForce[main_data$gender == 'F'])


M_mean_SBF <- mean(main_data$sit.and.bend.forward_cm[main_data$gender == 'M'])
F_mean_SBF <- mean(main_data$sit.and.bend.forward_cm[main_data$gender == 'F'])

M_mean_SU <- mean(main_data$sit.ups.counts[main_data$gender == 'M'])
F_mean_SU <- mean(main_data$sit.ups.counts[main_data$gender == 'F'])

M_mean_BrJ <- mean(main_data$broad.jump_cm[main_data$gender == 'M'])
F_mean_BrJ <- mean(main_data$broad.jump_cm[main_data$gender == 'F'])


data1 <- data.frame(Category = c('Male','Female'), Value = c(M_mean_grip,F_mean_grip))
data2 <- data.frame(Category = c('Male','Female'), Value = c(M_mean_SBF,F_mean_SBF))
data3 <- data.frame(Category = c('Male','Female'), Value = c(M_mean_SU, F_mean_SU))
data4 <- data.frame(Category = c('Male','Female'), Value = c(M_mean_BrJ, F_mean_BrJ))

color_palette <- c("red", "green", "blue", "purple")


p1 <- ggplot(data1, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Gender", y = "Mean Grip Force(kg)",cex.main=15)+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Grip Force ")
  # scale_y_continuous(breaks=seq(0:50:5))

p2 <- ggplot(data2, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Gender", y = "Mean Sit and Bend forward(cm)")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Sit and Bend forward")

p3 <- ggplot(data3, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Gender", y = "Mean Situp Count")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Situp Count")

p4 <- ggplot(data4, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Gender", y = "Mean Broad Jump(cm)")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Broad Jump")

grid.arrange(p1, p2, p3, p4, ncol = 2)

```

-Only In 'Sit and Bend Forward' Bar plot we can see mean performance of Female is higher than mean performance of Male

## \<\>. Which Gender has better 'Body attributes' ?

```{r}
################################################################################
# mean of PA and gender
################################################################################

M_mean_H <- mean(main_data$height_cm[main_data$gender == 'M'])
F_mean_H <- mean(main_data$height_cm[main_data$gender == 'F'])


M_mean_W <- mean(main_data$weight_kg[main_data$gender == 'M'])
F_mean_W <- mean(main_data$weight_kg[main_data$gender == 'F'])

M_mean_F <- mean(main_data$body.fat_.[main_data$gender == 'M'])
F_mean_F <- mean(main_data$body.fat_.[main_data$gender == 'F'])




data1 <- data.frame(Category =c('Male','Female'), Value = c(M_mean_H,F_mean_H))
data2 <- data.frame(Category =c('Male','Female'), Value = c(M_mean_W,F_mean_W))
data3 <- data.frame(Category =c('Male','Female'), Value = c(M_mean_F, F_mean_F))

color_palette <- c("red", "green", "blue", "purple")


p1 <- ggplot(data1, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Gender", y = "Mean Height(cm)",cex.main=15)+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Height ")
  # scale_y_continuous(breaks=seq(0:50:5))

p2 <- ggplot(data2, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Gender", y = "Mean Weight(kg)")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Weight")

p3 <- ggplot(data3, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Gender", y = "Mean Body Fat(%)")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Body Fat")



grid.arrange(p1, p2, p3, ncol = 2)

```

-Mean Body Fat percentage is higher in Female than Male

-Mean Height is lower in Female than Male

-Mean Weight is lower in Female than Male

-Chance of male performing better than female is high

## \<\>. Which class has better performing people?

```{r}
################################################################################
# mean of PA and class 
################################################################################

A_mean_grip <- mean(main_data$gripForce[main_data$class == 'A'])
B_mean_grip <- mean(main_data$gripForce[main_data$class == 'B'])
C_mean_grip <- mean(main_data$gripForce[main_data$class == 'C'])
D_mean_grip <- mean(main_data$gripForce[main_data$class == 'D'])

A_mean_SBF <- mean(main_data$sit.and.bend.forward_cm[main_data$class == 'A'])
B_mean_SBF <- mean(main_data$sit.and.bend.forward_cm[main_data$class == 'B'])
C_mean_SBF <- mean(main_data$sit.and.bend.forward_cm[main_data$class == 'C'])
D_mean_SBF <- mean(main_data$sit.and.bend.forward_cm[main_data$class == 'D'])

A_mean_SU <- mean(main_data$sit.ups.counts[main_data$class == 'A'])
B_mean_SU <- mean(main_data$sit.ups.counts[main_data$class == 'B'])
C_mean_SU <- mean(main_data$sit.ups.counts[main_data$class == 'C'])
D_mean_SU <- mean(main_data$sit.ups.counts[main_data$class == 'D'])

A_mean_BrJ <- mean(main_data$broad.jump_cm[main_data$class == 'A'])
B_mean_BrJ <- mean(main_data$broad.jump_cm[main_data$class == 'B'])
C_mean_BrJ <- mean(main_data$broad.jump_cm[main_data$class == 'C'])
D_mean_BrJ <- mean(main_data$broad.jump_cm[main_data$class == 'D'])


data1 <- data.frame(Category = c('A','B','C','D'), Value = c(A_mean_grip,B_mean_grip,C_mean_grip,D_mean_grip))
data2 <- data.frame(Category = c('A','B','C','D'), Value = c(A_mean_SBF, B_mean_SBF, C_mean_SBF,D_mean_SBF))
data3 <- data.frame(Category = c('A','B','C','D'), Value = c(A_mean_SU, B_mean_SU, C_mean_SU,D_mean_SU))
data4 <- data.frame(Category = c('A','B','C','D'), Value = c(A_mean_BrJ, B_mean_BrJ, C_mean_BrJ,D_mean_BrJ))

# color_palette <- c("red", "green", "blue", "purple")


p1 <- ggplot(data1, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Class", y = "Mean Grip Force(kg)",cex.main=15)+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Grip Force ")
  # scale_y_continuous(breaks=seq(0:50:5))

p2 <- ggplot(data2, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Class", y = "Mean Sit and Bend forward(cm)")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Seat and Bend forward")

p3 <- ggplot(data3, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Class", y = "Mean Situp Count")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Situp Count")

p4 <- ggplot(data4, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Class", y = "Mean Broad Jump(cm)")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Broad Jump")

grid.arrange(p1, p2, p3, p4, ncol = 2)

```

CONCLUSION: people of age group 25-44 are considered most Fit if they have

-GF above 38.62kg,

-SBF above 21.39cm

-Sit up count above 47

-Br J above 202.4

then he or she surely belongs to A

## \<\>. How body attributes effects the individual being in a group?

```{r}
################################################################################
# Box plots of Body attributes and age
################################################################################


data <- data.frame(
  Category = main_data$class,
  Value = main_data$height_cm
)
bp1 <- ggplot(data, aes(x = Category, y = Value, fill = Category)) +
  geom_boxplot(width = 0.2) +
  scale_fill_manual(values = c("orange", "green", "skyblue", "purple")) +
  coord_flip()+
  # theme_minimal() +
  labs(title = "Box Plots Between Class and Height",x='Class',y='Height(cm)')
# print(bp1)

data <- data.frame(
  Category = main_data$class,
  Value = main_data$weight_kg
)
bp2 <- ggplot(data, aes(x = Category, y = Value, fill = Category)) +
  geom_boxplot(width = 0.2) +
  scale_fill_manual(values = c("orange", "green", "skyblue", "purple")) +
  coord_flip()+
  # theme_minimal() +
  labs(title = "Box Plots Between Class and weight",x='Class',y='weight(kg)')
# print(bp2)

data <- data.frame(
  Category = main_data$class,
  Value = main_data$body.fat_.
)
bp3 <- ggplot(data, aes(x = Category, y = Value, fill = Category)) +
  geom_boxplot(width = 0.2) +
  scale_fill_manual(values = c("orange", "green", "skyblue", "purple")) +
  coord_flip()+
  # theme_minimal() +
  labs(title = "Box Plots Between Class and Body Fat%",x='Class',y='Body Fat%')
# print(bp3)



grid.arrange(bp1, bp2,bp3, ncol = 2)






################################################################################
# Bar plots of mean BA and class 
################################################################################

A_mean_H <- mean(main_data$height_cm[main_data$class == 'A'])
B_mean_H <- mean(main_data$height_cm[main_data$class == 'B'])
C_mean_H <- mean(main_data$height_cm[main_data$class == 'C'])
D_mean_H <- mean(main_data$height_cm[main_data$class == 'D'])

A_mean_W <- mean(main_data$weight_kg[main_data$class == 'A'])
B_mean_W <- mean(main_data$weight_kg[main_data$class == 'B'])
C_mean_W <- mean(main_data$weight_kg[main_data$class == 'C'])
D_mean_W <- mean(main_data$weight_kg[main_data$class == 'D'])

A_mean_F <- mean(main_data$body.fat_.[main_data$class == 'A'])
B_mean_F <- mean(main_data$body.fat_.[main_data$class == 'B'])
C_mean_F <- mean(main_data$body.fat_.[main_data$class == 'C'])
D_mean_F <- mean(main_data$body.fat_.[main_data$class == 'D'])




data11 <- data.frame(Category = c('A','B','C','D'), Value = c(A_mean_H,B_mean_H,C_mean_H,D_mean_H))
data21 <- data.frame(Category = c('A','B','C','D'), Value = c(A_mean_W, B_mean_W, C_mean_W,D_mean_W))
data31 <- data.frame(Category = c('A','B','C','D'), Value = c(A_mean_F, B_mean_F, C_mean_F,D_mean_F))

color_palette <- c("red", "green", "blue", "purple")


p11 <- ggplot(data11, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Class", y = "Mean Height(cm)")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Height ")
  # scale_y_continuous(breaks=seq(0:50:5))

p21 <- ggplot(data21, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  labs(title = "", x = "Class", y = "Mean Weight(kg)")+
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Weight")

p31 <- ggplot(data31, aes(x = Category, y = Value, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label =round( Value,2)), vjust = -0.2, size = 3) +
  labs(title = "", x = "Class", y = "Mean Body Fat(%)")+
  # scale_fill_manual(values = color_palette) +
  labs(title = "Mean Body Fat")


grid.arrange(p11, p21, p31, ncol = 2)

```

-as all class have more or less same distribution of age so people in age group 25-44 are most fit

who has height above 16.87 cm and

weight below 64.42 and

body fat % below 20.54% .

so if an individual surely wants to be in class A he or she has to satisfy there Body attributes,

## \<\> Is there any relation between Height and Broad jump?

```{r}

data <- data.frame(
  X = main_data$height_cm,
  Y = main_data$broad.jump_cm
)


scatter_plot <- ggplot(data, aes(x = X, y = Y)) +
  geom_point(size=0.5) +
  labs(title = "Scatter Plot b/w Height and Broad Jump", x = "Height(cm)", y = "Broad Jump(cm)")


print(scatter_plot)

correlation <- cor(main_data$height_cm, main_data$broad.jump_cm,method = "kendall")
cat("Correlation: ",correlation)
```

-From Scatter Plot as height increases Amount of broad jump measure also increases

-so they are +co related

-Also from kendall Tau's correlation measurement we get +ve correlation

## \<\>What is the relation between Body Attributes and Physical Attributes?

```{r}
################################################################################
# Correlation matrix
################################################################################

num_data<-main_data[sapply(main_data,is.numeric)]
cor_matrix <- cor(num_data)
corrplot(
  cor_matrix,
  method = "color",        
  diag = FALSE,
  type = "full",           
  tl.cex = 0.8,            
  tl.col = "black",         
  tl.srt = 45,             
  tl.pos = "t",            
  addCoef.col = "black",   
  col = colorRampPalette(c("red","white", "green"))(100),  
  number.cex = 0,         
  mar = c(4, 0, 0.1, 0)       
)
title("Correlation Matrix(Heat Map) Plot", line = 2, outer = FALSE)


```

-age is negatively corelatd to all Physical attributes

-height is positively correlated to sit_ups_counts and broad_jump,weight

-height is negatively correlated to body_fat_percentage

-body_fat_percentage is negatively correlated to sit_ups_count and broad_jump
